\section{Background}
\label{sec:background}

\paragraph{Coherence vs.\ consistency.}
Cache coherence is a per-location property. It ensures that one address
has a single coherent history across private caches. Memory consistency
governs what a program may observe across multiple locations
~\cite{AdveGharachorloo1996,SorinHillWood2020Primer}. The distinction is
central here. A coherent system can still expose a mixed multi-field
record if the producer writes fields one at a time and the reader races
with that update. On weakly ordered architectures such as ARMv8, this
problem is more visible because inter-field ordering must be imposed
explicitly with release, acquire, and barrier operations
~\cite{Pulte2018SimplifyingARM,Flur2016ARMv8,Maranget2012Tutorial,%
ARMv8ARM}.

\paragraph{Supervisory monitoring in \CPS.}
Commercial safety-oriented SoCs increasingly provide a supervisory core
or safety island alongside the main application complex
~\cite{ArmAutoSolutionsOverview,NVIDIADriveFSI}. In mixed-criticality
terms, the question is not only where the high-criticality task runs but
also how it obtains a predictable and trustworthy view of the
low-criticality world~\cite{BurnsDavis2022MixedCrit,Kopetz2022RealTime}.
This paper focuses on that interface.

\paragraph{Runtime monitoring and publication primitives.}
Signal temporal logic provides a natural language for bounded-window
monitoring over sampled signals, and robust online monitoring is now
well understood~\cite{MalerNickovic2004,FainekosPappas2009,%
Deshmukh2017,Bartocci2018CPSMonitoring}. The limiting issue in this
study is therefore not the monitor formalism, but how the signal reaches
the monitor. Sequence locks and double buffers are standard ways to
publish multi-field state safely, while predictable-coherence work
motivates comparing direct sharing against an explicit-transfer path
~\cite{LamportSeqLock1977,Boehm2012Seqlocks,McKenney2019RCU,%
Hassan2017PredictableCoherence,Hassan2022IntegratingCOTSCoherence}.
RCU is therefore relevant background, but we do not include it in the
matrix because grace-period machinery addresses memory lifetime and
eventual quiescence rather than the bounded-window freshness contract
studied here.

\paragraph{Direct sharing vs.\ explicit transfer.}
Those two transports expose different architectural tradeoffs. Direct
sharing keeps one active record in the coherent hierarchy and lets the
monitor observe producer updates without an extra copy. An
explicit-transfer design decouples the monitor from the producer's
active cache line, but it pays a transfer path, a second memory region,
and its own publication problem on the mirror side. The key point for
this paper is that transport and publication discipline should not be
conflated. A DMA mirror with no record-level handshake is still
incorrect, just in a different place.
