\subsection{Experimental Platform and Metrics}
\label{sec:eval-method}

\paragraph{Simulation platform.}
All experiments run on gem5~25.1.0.0 in ARM SE mode with the
\texttt{TimingSimpleCPU} front end and the Ruby
\texttt{MESI\_Two\_Level} coherence protocol
~\cite{Binkert2011gem5,LowePower2020gem5,SorinHillWood2020Primer}. A
three-CPU system hosts one producer thread, one monitor thread, and one
joiner thread. Cache contention is added with 0, 2, or 4 additional
stressor CPUs running disjoint-line thrashers. For the two DMA-mirror
architectures we attach a custom pull engine to the Ruby DMA sequencer.
The setup is not a cycle-accurate model of a production AMBA CHI fabric.
It is an architecture-level testbed with direct visibility into Ruby
message counts and directory traffic.

\paragraph{Interference model.}
The 0/2/4-stressor sweep is a synthetic proxy for increasing co-runner
pressure on shared caches and directory resources. The stressors touch
disjoint lines, so they inject hierarchy interference without creating
false sharing on the witness region itself. The goal is not to model one
named application mix, but to test whether the publication ranking
persists as background pressure rises.

\paragraph{Workloads.}
Each cell processes 5000 publications against two workload variants. The
first, \emph{captured / periodic\_suppression}, is a sparse-divergence
trace in which realized behavior matches intent except during short
bursts. The second, \emph{synthesized / duty\_bias}, introduces a
persistent +15\% multiplicative bias between intended and realized duty
values, clipped to the valid duty range. With $\varepsilon=0.1$ and an
\STL window of 8 samples, this sustained-bias workload produces
repeated monitor events, ranging from 180 to 949 \STL violations per
5000-publication run across the correct designs. The matrix therefore
spans 5 architectures, 2 workloads, and 3 contention levels, for 30
cells total.

\paragraph{Benchmark structure.}
The producer publishes one 64-byte record per trace step. For the two
DMA-mirror architectures, a custom pull engine copies one 64-byte record
from producer memory into the mirror every 1000 ns, which is the fixed
transfer cadence for the study. The monitor reads as fast as the
publication protocol allows and aligns successful snapshots to the
evidence stream by epoch. This design intentionally stresses the
publication path rather than emulating a wall-clock control loop.

\paragraph{Metrics.}
We report five primary metrics derived directly from Ruby statistics and
bench-side counters.
\begin{itemize}
  \item \textbf{Torn-read fraction:} the fraction of attempted monitor
    reads flagged as mixed records.
  \item \textbf{CPU-side coherence messages per publish:} the sum of
    L1-to-directory and directory-to-memory traffic, normalized by the
    number of publishes.
  \item \textbf{DMA messages per publish:} the
    \texttt{dma\_cntrl0.requestToDir} count per publish, which is nonzero
    only for the DMA-based architectures.
  \item \textbf{Protocol retries per run:} the number of read attempts
    rejected by the publication primitive and retried instead of being
    returned as data.
  \item \textbf{Whole-workload simulated time:} the simulated time
    between \texttt{m5\_reset\_stats} and \texttt{bench\_dump\_stats}.
\end{itemize}
Because the producer runs unpaced, the whole-workload time is a
throughput-style proxy for the cost of retiring the workload under each
architecture. It is not a direct publication-to-alarm latency bound.

\paragraph{Interpretation of the DMA comparison.}
The DMA-based designs are not intended to represent a fully
non-coherent monitor subsystem. The mirror still resides in the same
coherent substrate. What changes is the communication pattern: the
monitor no longer reads the producer's active record directly, but
instead relies on an explicit pull into a second memory region. The
comparison therefore answers a narrower and cleaner architecture
question: when the substrate is coherent in both cases, is it cheaper
and simpler to publish directly with a disciplined primitive, or to pay
an explicit transfer path and validate a mirror?
